\documentclass[sigconf]{acmart}

\input{format/i523}

\begin{document}
\title{Big Data in Sports Visualization}


\author{Josh Lipe-Melton}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Indiana University}
  \streetaddress{4400 E Sheffield Dr}
  \city{Bloomington} 
  \state{Indiana} 
  \postcode{47408}
}
\email{jlipemel@umail.iu.edu}

\renewcommand{\shortauthors}{J. Lipe-Melton}


\begin{abstract}
We discuss Big Data Analytics in ''other sports'' (not baseball) and how it is being used to improve and evaluate performance via visualization of 
\end{abstract}

\keywords{sports, analytics, data visualization, spatial rendering, HID105, I523}


\maketitle




\section{Introduction} 

Data visualization in sports other than baseball is a rapidly growing field due to the explosion in amount of data captured about these sports. A general problem is that it is very difficult to quantify most team sports because of the chaotic nature of the placement of the ball and of players. Furthermore, it is difficult to make use of massive amounts of data for real sports analysis. Big data allows vast quantities of data points to be assessed in a small amount of time. One example of a use case is SoccerStories, a data visualization tool for soccer which makes use of several different techniques to represent real world events in a series of diagrams. Snapshot is a similar example that uses complex formulas to turn huge quantities of data points into easily readable representations using graphs, symbols, and heat maps. Director’s Cut is another program that enhances simple statistics with more useful context. There are various methods and tools for sports data visualization.

\section{Shot Chart Visualization}

Team sport data analysts are met with the challenge of the dynamic, and even chaotic position of players and the ball or puck. Data visualization, however, allows decisions to be made based on large data sets that would otherwise be difficult to understand. Snap Shot is an example of a hockey data visualization technique that allows teams to identify the position, trajectory, and effectiveness of shots throughout the course of a game or season \cite{SnapShot}. A specific use case was to analyze several theories about ''sweet spots'' on the ice, or positions that a shooter is more likely to score from. One such theory is that goalies tend to be right handed, and that makes it more difficult for them to block shots from their right side due to holding their stick in that hand \cite{SnapShot}. This is an intuition held by many high level coaches that was proved to be false after numerous queries through SnapShot \cite{SnapShot}. Similarly, NBA shooting was modeled by experts at MIT's sloan sports analytics conference \cite{CourtVision}. Using color and size to represent different data points for shots taken layered onto different positions on the court, one can quickly see where players tend to shoot from and where they are most effective. This type of analysis could be done to show coaches and players where they should be shooting from or what types of plays to draw up for which players.

\subsection{Basketball Visualization}

Shooting is not the only use of spatial relationships in basketball. By associating spatial data with statistics such as fouls, defensive effectiveness, or rebounds, other types insights can be made. For example, if an opposing player is in foul trouble, a coach can draw up a play to isolate that player in the position where they are most likely to foul again through use of a visualization tool. Furthermore, by layering different spatial visualizations on top of one another, more insights can be made. ''By overlaying individual players' shot charts with team defense charts we can predict which players in the league will perform best and worst against any team defense.'' \cite{CourtVision) Although CourtVision can not effectively do this right now, it contests that ''the combined abilities of spatial analysis and visual analytics.. enable us to answer questions, but perhaps more importantly, they enable us to communicate these concepts to diverse audiences; almost anyone can understand a well-designed map or chart.''\cite{CourtVision}

\section{Tennis}

Tennis analysis is also being transformed by the massive influx of data into sports, and all this data is represented visually in a variety of ways. Serves can be represented by lines to identify patterns in placement and height from an opponent's serve. Moreover, by associating serves with whether the point was won or lost can identify strengths or weaknesses from areas in service. Andy Murray, for example, used a tool called Hawk-Eye to analyze a loss in a tournament finals against his opponent, Kei Nikishori. Using this tool, he realized that almost every time he left a serve short he lost the point. As Game Set Map observes, ''if you're serving short 2nd serves to Nikishori at important points than Nikishori is going to be all over the return and you'll be playing catch up all game!''\cite{GameSetMap} Using this observation, Murray pushed his 2nd serve farther back in his next matchup with Nikishori and ultimately winning the match. While these types of changes may seem insignificant, players and coaches alike place high value on these sorts of insights. Andy Murray is just one example of a frequent user of Hawk Eye. In fact, Murray took it a step too far when he was caught filming a practice session against an opponent without his permission.\cite{GameSetMap} Andy Murray clearly places great value on the data and visualization techniques used in tools such as Hawk Eye.

\section{Soccer}

Similar to hockey or basketball, soccer is a difficult sport to quantify due to the seemingly subjective methods of evaluating different plays and the randomness of the locations at which plays start and end. One of the most basic and commonly employed data visualization techniques for soccer games is a timeline \cite{SoccerStories}. This method takes advantage of a predetermined statistic in the game: the length of each half. By using symbols to represent real world events, such as a ball representing a goal, and placing these symbols along the timeline, one is able to gain a limited understanding of the events in the game. This might give simple statistics such as shots, possession, or fouls, more context within the flow of the game and allow a user to gain more insight into what is happening. 

\subsection{Field Position Identification}

Deeper analysis of a soccer game typically relates a game event such as a shot or pass with positions on the field. According to soccerstories, ''the soccer field is the primary object of observation and analysis in soccer. Analysts construct their mental model over the spatial arrangement of the team, and its motion, over time.''\cite{SoccerStories} This type of data can be accrued through the use of wearable technology or through video \cite{SoccerStories}. One common method of this visualizing this data is a ''heatmap, through which player's most frequent positions is displayed by density'' \cite{SoccerStories}. A heat map allows for an intuitive and instantaneous evaluation of a player’s positioning which might otherwise take the length of a match to evaluate. This technique could also be used to identify a player that does not run back to play defense or a player that gets pulled out of position easily. By identifying visual patterns to make insights such as these, teams can gain competitive advantages \cite{SoccerStories}.

\subsection{Set Piece Analysis}

An important part of the scouting report for a soccer game is corner kicks and free kicks. According to researchgate, about 30 percent of goals in soccer come from these situations \cite{ResearchGate}. Identifying patterns in where a team likes to direct set pieces allows a coach to set up their team in an advantageous position. For example, by using a heat map to show the frequency and effectiveness of crosses to different positions from these situations allows coaches to put a zonal marker in position to neutralize the threat. Furthermore, this method can cut back on time spent analyzing video or emphasize insights gained from watching video. By identifying which free kicks or corner kicks were passed to a player very close to the initial position of the ball, a coach or player can quickly prepare themselves for any ''trick plays'' an opponent has used in the past.

\subsection{Flow Graph Uses}

Another method of relating simple statistics to locations on the field is a flow graph, ''where the size of the nodes shows player's role in the game and the links show the connections between players'' \cite{SoccerStories}. A flow graph relates simple statistics about individual players to other individual players and complex insights to be made quickly, such as discerning which players like to pass to one another or which player has a greater impact on the game \cite{SoccerStories}. This could also be useful in evaluating a team’s tendencies such as identifying an inability to attack down the right side or give up more shots on the left. These types of tendencies can be used to make decisions such as what formation to play or what spaces to run into on counterattacks.

\section{Team Shape Analysis}

There are several ways team based analysis can often be utilized in data visualization. Director’s Cut creates an analysis of a team’s ''back four,'' or defensive line, which can be portrayed by simply drawing a line connecting each of the defenders \cite{directorscut}. A team’s coach can use this line to identify situations where the back line maintains good flat shape defensively in order to play an offside trap, or alternately situations where the defense gets stretched and could allow a player to get behind the defense \cite{directorscut}.

\subsection{Player to Player Spatial Relationships}

A player’s proximity to opposing players is another factor that could be useful in evaluating performance. Director’s cut, for example, breaks down player proximity into three separate categories: no pressure, weak pressure, and strong pressure \cite{directorscut}. This is done by segmenting the soccer field into one meter by one meter squares \cite{directorscut}. Each square is then assigned several attributes regarding the closest player and their speed, direction, and proximity to the square \cite{directorscut}. ''Pressure'' on each player can then be determined by viewing the attributes assigned to the square he or she occupies \cite{directorscut}. This can be extremely helpful in analysis of an individual player’s ability to cope with situations that tend to force mistakes \cite{directorscut}. For example, a player’s pass completion percentage, the number of passes they complete divided by the number of passes they attempt, can be a useful but slightly misleading statistic. A forward will tend to have lower pass completion percentage than a defender, for example, due to the closer proximity of the other team’s defenders. By associating pass completion with no, weak, or strong pressure, however, individual players can be analyzed and compared to one another more easily. Furthermore, a player that misses more passes under no pressure can be shown what they are doing wrong and correct the problem very quickly. By analyzing what choices lead to passes that end up being cheap giveaways, a player could quickly improve on an important aspect of their game. A player missing passes under no pressure is often likely to be making bad decisions more than lacking talent or technique. By displaying what areas of the field and what distances that poor passes are made over, a player can be better informed and make better decisions on the field. Visualization, in this case, could rapidly accellerate the transition from data to decisions.

\section{Conclusions}

In conclusion, data visualization is a widely used application of big data in sports. Technology such as player trackers and video analysis allows players, coaches, and managers to gather and communicate insights into sports performances. Many of these visualization techniques are combinations of simple statistics and context such as field position or time remaining. Basketball and hockey use data visualization for analysis of high percentage shots, while soccer focuses on tactics, formations, and player evaluations. There are numerous tools for each of these, all with various methods of carrying out these tasks.

\section{Acknowledgments}

I would like to thank Gregor von Laszewski for generously fixing compile errors and Juliette Zerick for providing accurate and specific feedback.

\bibliographystyle{ACM-Reference-Format}
\bibliography{report} 

\end{document}
